---
title: "Statistical Distribution Task1"
author: "Innocent T. Novart"
date: "2022-11-11"
output: html_document
---

# Task 1 Question:

## A company is considering automating the ordering system. The estimate indicates that the investment will be economically justified if the number of orders exceeds 200 a day. The average daily number of orders in a company with the same business profile within 35 days was 225, with a standard deviation of 32. Determine 95% and 99% confidence interval in which the average number of orders will be found. What decision should the authorities of this company take?


### Solution For 95% confidence level

```{r}
n = 35     #  since the sample size is greater than 30, then we shall use the normal distribution.
xbar = 225
sigma = 32
cl = 0.95
alpha = 1-cl
z = qnorm(1-alpha/2)
z
EBM = z*(sigma/sqrt(n))
EBM
lower_end = xbar - EBM
upper_end = xbar + EBM
ci_0.95 = c(lower_end,upper_end)
ci_0.95

#  With the probability of 95%, the average number of orders will be found between 214 and 235 orders per day.
```


### Solution for 99% confidence level

```{r}
sigma = 32
xbar = 225
n = 35
cl = 0.99
alpha = 1-cl
z = qnorm(1-alpha/2)
z
EBM = z*(sigma/sqrt(n))
EBM
lower_end = xbar - EBM
upper_end = xbar + EBM
ci_0.99 = c(lower_end,upper_end)
ci_0.99

#  With the probability of 99%, the average number of orders will be found between 211 and 239 orders per day.
```


## What decision should the authorities of this company take?

## Answer

### From the results, it has been statistically discovered that the average daily orders from a similar business profile are found to be between 214 and 235, and  between 211 and 239 orders per day at 95% and 99% confidence levels respectively. Since both probabilities have given an interval greater than 200 orders per day, the investment is economically justifiable. Therefore, the company should automate the ordering system.


