---
title: "Hypothesis Testing of Proportion"
author: "Innocent Novart"
date: "2022-11-10"
output: html_document
---

## Question: Suppose 60% of citizens voted in last election. 85 out of 148 people in a telephone survey said that they voted in current election. At .05 significance level, can we reject the null hypothesis that the proportion of voters in the population is above 60% this year?

```{r}
# Since the % of sample proportion (85/148*100 = 57%) is less than the hypothesized % proportion (60%), then this is a lower tail test.

pbar = 85/148    #sample proportion
p0 = 0.6         #hypothesized proportion
n = 148          #sample size
alpha = 0.05     #significance level
z = (pbar-p0)/sqrt(p0*(1-p0)/n)     #test statistic
z
z.alpha = qnorm(1-alpha)
-z.alpha                            #critical element

# From the results, the test statistic (z = -0.64) is not within the critical area. Therefore, 5% significance level, we fail to reject the null hypothesis that the proportion of voters in the population is above 60%.

pval = pnorm(z)
pval

# From the result, the p-value (26%) is greater than the significance level. Therefore, we fail to reject the null hypothesis that the proportion of voters in the population is above 60%.
```

## Question: Suppose that 12% of apples harvested in an orchard last year was rotten. 30 out of 214 apples in a harvest sample this year turns out to be rotten. At .05 significance level, can we reject the null hypothesis that the proportion of rotten apples in harvest stays below 12% this year?

```{r}
# Since the % of sample proportion (30/214*100 = 14%) is greater than the hypothesized % proportion (12%), then this is an upper tail test.

pbar = 30/214
p0 = 0.12
n = 214
alpha = 0.05
z = (pbar-p0)/sqrt(p0*(1-p0)/n)
z
z.alpha = qnorm(1-alpha)
z.alpha

# From the results, the test statistic (z = 0.91) is not within the critical area. Therefore, at 5% significnce level, we fail to reject the null hypothesis that the proportion of rotten apples in harvest stays below 12%.

pval = pnorm(z, lower.tail = FALSE)
pval

# From the result, the p-value (18%) is greater than the significance level. Therefore, we fail to reject the null hypothesis that the proportion of rotten apples in harvest stays below 12%.
```

## Question: Suppose a coin toss turns up 12 heads out of 20 trials. At .05 significance level, can one reject the null hypothesis that the coin toss is fair?

```{r}
pbar = 12/20
p0 = 0.5
n = 20
alpha = 0.05
z = (pbar-p0)/sqrt(p0*(1-p0)/n)
z
z.half.alpha = qnorm(1-alpha/2)
c(-z.half.alpha,z.half.alpha)

# From the results, the test statistic (z = 0.89) is not within the critical area. Therefore, at 5% significnce level, we fail to reject the null hypothesis that the coin toss is fair.


```

